How to Avoid OOTA Without Really Trying

### 2024-09-10 paper/ootavacuous.tex
ootavacuous: Consolidate abstract and introduction
No point in saying the same thing twice, and especially not in the same
column of the same page.

Signed-off-by: Paul E. McKenney <paulmck@kernel.org>

\begin{abstract}
	The out-of-thin-air (OOTA) properties
	of \co{memory_order_relaxed} in the C++ memory model
	have caused considerable consternation over the years.
	Attempts to create memory models that rule out OOTA behaviors
	have been either non-executable, complex, or unloved by C++
	implementers.
	But at the same time, we know of no instances of OOTA behavior
	in real C++ implementations.

	We focus on shared-memory programs and
	C++ implementations based on traditional compilers and
	computing hardware, including CPUs and GPGPUs.
	This context permits us to consider the detailed relations
	between source code and machine code required by the restricted
	nature of volatile atomic accesses.
	The OOTA problem and the related challenge of coming to grips
	with semantic dependency are much more tractable at the hardware
	level than at the source level, thanks to existing formal
	hardware models.
	We show that these models' constraints prevent OOTA cycles
	from occurring in undefined-behavior-free C++ programs running
	on compiler-based implementations, provided the cycles involve
	only volatile atomics.
	We also extend this work to nonvolatile atomics by defining
	``quasi-volatile'' behavior that we expect C++ implementations
	will adhere to if they confine themselves to single-thread analysis.
\end{abstract}
\maketitle{}

@@

This paper shows that compiler-based C++ implementations subject to
reasonable constraints on how they treat accesses to atomic objects
cannot exhibit out-of-thin-air (OOTA) cycles.
It follows that these implementations need to take almost no special
actions to avoid OOTA.
In fact, for many compilers the constraints boil down to a simple ``Don't
invent or duplicate atomic accesses'', which the compiler should not
do anyway if it is to evaluate arithmetic expressions involving atomic
accesses in accordance with the abstract machine.
Absolutely no changes are required to either the standard or to user code.

>>>>

\begin{abstract}
	Attempts to create memory models that forbid out-of-thin-air
	(OOTA) behaviors of C++ \co{memory_order_relaxed} accesses
	have been either non-executable, complex, or unloved by
	implementers.
	However, we know of no instances of OOTA behavior in real C++
	implementations.

	We focus on shared-memory programs and C++ implementations based
	on traditional compilers and hardware, including CPUs and GPGPUs,
	thus permitting easier analysis of semantic dependencies and
	OOTA behaviors, thanks to existing formal hardware models.
	We show that these models' constraints prevent OOTA cycles from
	occurring in undefined-behavior-free C++ programs running on
	compiler-based implementations, provided the cycles involve only
	volatile atomics.
	We also extend this work to nonvolatile atomics by defining
	``quasi-volatile'' behavior that we C++ implementations
	must adhere to in order to evaluate arithmetic expresssions
	involving atomic accesses in accordance with the abstract
	machine.

	It follows that such implementations need take no special action
	to avoid OOTA, and that no changes are required to user code or
	to the standard.
\end{abstract}
\maketitle{}

@@
